SPENCER THESIS - WORK IN PROGRESS

### 2016-05-25 doc/thesis_swe.tex
Updated the correlation function and smooth-component correction sections.

%%==========================================================================================================================================
\section{Cosmological Background}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Dark Matter Halos}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Filaments, Voids, and the Smooth-Component Correction}

@@

\noindent Combining this result with Equation \eqref{temp_solution} leads to the desired proper volume element given by
\begin{equation}
dV_p=\frac{d_Hd_M(z)^2}{E(z)(1+z)^3}dzd\Omega.
\end{equation}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Gravitational Lensing}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Galaxies as Ellipses}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Correlation Functions}

A common tool used by cosmologists to measure...

@@

The mathematical tool that measures this average correlation as a function of separation distance between two galaxies is called the ellipticity-ellipticity ($\varepsilon$-$\varepsilon$) correlation function. A quick overview of this function is given below, but readers that are unfamiliar with correlation functions in the context of cosmology or want a visual aid should see Appendix (??).

Given two galaxies with ellipticities $\varepsilon_i$ and $\varepsilon_j$ whose orientation is offset by polar angle $\alpha$, the tangential and cross-component components of the ellipticities, $\varepsilon_t$ and $\varepsilon_\times$ respectively, for each are defined as
\begin{align}
\varepsilon_{n_t}&=-\operatorname{Re}\left(\varepsilon e^{-2i\alpha}\right),\\
\varepsilon_{n_\times}&=-\operatorname{Im}\left(\varepsilon e^{-2i\alpha}\right).
\end{align}

\noindent Instead of using the autocorrelation and cross-correlation between these quantities, it turns out to be more convenient to define the combinations

\begin{align}\label{gg_corr_def}
\xi_\pm&=\left<\varepsilon_{i_t}\varepsilon_{j_t}\right>\pm\left<\varepsilon_{i_\times}\varepsilon_{j_\times}\right>\\
\xi_\times&=\left<\varepsilon_{i_t}\varepsilon_{j_\times}\right>
\end{align}

\noindent where $\theta$ is the position vector of the first galaxy and $\Delta\vec{\theta}$ is the angular separation vector between the pair of galaxies \cite{schneider}. Gravitational lensing shears ellipticities tangentially, so $\xi_\times(\Delta\theta)$ should vanish with enough galaxies. (\textcolor{red}{Why is $\xi_-$ expected to be zero?}), which leaves $\xi_+(\Delta\theta)$ as the desired probe of lensing mass.

@@

%%==========================================================================================================================================
\section{Pangloss: Line of Sight Mass Reconstructions}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Model Assumptions}
While \texttt{Pangloss} may be used more generally, the present analysis makes some additional strong assumptions to simplify the problem for a first attempt at making weak lensing predictions.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{The Millennium Simulation}
\texttt{Pangloss} cannot be used to make dark matter mass maps using existing galaxy catalogs until it is tested on a simulated universe with known dark matter structure to determine how accurately and precisely it predicts the lensing of background sources. For this purpose, \texttt{Pangloss} was tested on galaxy catalogs from the Millennium Simulation, an N-body simulation consisting of over 10 billion dark matter `particles' each representing a billion solar masses and populated with about 20 million galaxies \cite{millennium_simulation}. The simulation contains baryonic and dark matter structure believed to be similar to our own universe, and uses cosmological parameters from WMAP 1\textsuperscript{st}-year data analysis which assumed a matter density of $\Omega_m=0.25$, a dark energy density of $\Omega_\Lambda=0.75$, and a Hubble constant of $H_0=73$ km$\,$s$^{-1}\,$Mpc$^{-1}$ (\textcolor{red}{CITE!!}). From the work of Hilbert et al. \cite{ray_tracing}, high resolution maps of the ray-traced shear and convergence of galaxy sources at $z=1.3857$ are publicly available. From these maps, an estimate of the `true' lensing of background galaxy ellipticities when traveling through the foreground mass of the Millennium Simulation can be calculated using Equation \eqref{lensing}.

@@

To calculate the convergence and shear contribution of each halo, the physical distance from the halo to the line of sight was needed. To increase the speed of distance calculations, the foreground halo redshifts were first binned to a grid of 100 equally-sized redshift bins from $z=0$ to the sourced redshift of $z=1.3857$ and then converted to physical distances using the cosmology defined by the Millennium Simulation.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Calculating the Smooth-Component Correction}

@@

(Introduce idea in Cosmological background, but cover image and specific details to implementation here.)\\

(Volume and Surface density plots for s.c.c. redshift slices)

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Surface and Volume densities for smooth-component correction redshift slices
% \begin{figure*}
%     \centering
%     \begin{subfigure}[H]{0.415\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/scc_vol_den.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{volume_densities}
%     \end{subfigure}
%     ~
%     %% Einstein Ring Figure
%     \begin{subfigure}[H]{0.425\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/scc_surf_deng.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{surface_densities}
%     \end{subfigure}
%     \caption{}
% \end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

>>>>

%%==========================================================================================================================================
\section{Cosmological Background} \label{background}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Dark Matter Halos} \label{halos}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Filaments, Voids, and the Smooth-Component Correction} \label{scc}

@@

\noindent Combining this result with Equation \eqref{temp_solution} leads to the desired proper volume element given by
\begin{equation}\label{pvolume_element}
dV_p=\frac{d_Hd_M(z)^2}{E(z)(1+z)^3}dzd\Omega.
\end{equation}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Gravitational Lensing} \label{grav_lensing}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Galaxies as Ellipses} \label{galaxies_as_ellipses}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Correlation Functions} \label{corr_functions}

@@

The mathematical tool that measures the average correlation as a function of separation distance between two galaxies is called the ellipticity-ellipticity ($\varepsilon$-$\varepsilon$) correlation function. A quick overview of this function is given below, but readers that are unfamiliar with correlation functions in the context of cosmology or want a visual aid should see Appendix (??).

Given two galaxies with ellipticities $\varepsilon_i$ and $\varepsilon_j$ whose orientation is offset by polar angle $\alpha$, the tangential and cross-component components of the ellipticities, $\varepsilon_t$ and $\varepsilon_\times$ respectively, for each are defined as
\begin{align}
\varepsilon_{k_t}&=-\operatorname{Re}\left(\varepsilon_k e^{-2i\alpha}\right),\\
\varepsilon_{k_\times}&=-\operatorname{Im}\left(\varepsilon_k e^{-2i\alpha}\right).
\end{align}

\noindent Instead of using the autocorrelation and cross-correlation of these quantities, it turns out to be more convenient to define the combinations

\begin{align}\label{gg_corr_def}
\xi_\pm(\Delta\theta)&=\left<\varepsilon_{i_t}(\vec{\theta})\varepsilon_{j_t}(\vec{\theta}+\vec{\Delta\theta})\right>\pm\left<\varepsilon_{i_\times}(\vec{\theta})\varepsilon_{j_\times}(\vec{\theta}+\vec{\Delta\theta})\right>\\
\xi_\times(\vec{\Delta\theta})&=\left<\varepsilon_{i_t}(\vec{\theta})\varepsilon_{j_\times}(\vec{\theta}+\vec{\Delta\theta})\right>
\end{align}

\noindent where $\vec{\theta}$ is the position vector of the first galaxy and $\Delta\vec{\theta}$ is the angular separation vector between the pair of galaxies \cite{schneider}. Gravitational lensing shears ellipticities tangentially, so $\xi_\times(\Delta\theta)$ should vanish with enough galaxies. (\textcolor{red}{Why is $\xi_-$ expected to be zero?}), which leaves $\xi_+(\Delta\theta)$ as the desired probe of lensing mass.

@@

%%==========================================================================================================================================
\section{Pangloss: Line of Sight Mass Reconstructions} \label{pangloss}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Model Assumptions} \label{assumptions}
While \texttt{Pangloss} may be used more generally, the present analysis makes some additional strong assumptions to simplify the problem for a first attempt at making weak lensing predictions.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{The Millennium Simulation} \label{ms}
\texttt{Pangloss} cannot be used to make dark matter mass maps using existing galaxy catalogs until it is tested on a simulated universe with known dark matter structure to determine how accurately and precisely it predicts the lensing of background sources. For this purpose, \texttt{Pangloss} was tested on galaxy catalogs from the Millennium Simulation, an N-body simulation consisting of over 10 billion dark matter `particles' each representing a billion solar masses and populated with about 20 million galaxies \cite{millennium_simulation}. The simulation contains baryonic and dark matter structure believed to be similar to our own universe, and uses cosmological parameters from WMAP 1\textsuperscript{st}-year data analysis which assumed a matter density of $\Omega_m=0.25$, a dark energy density of $\Omega_\Lambda=0.75$, and a Hubble constant of $H_0=73$ km$\,$s$^{-1}\,$Mpc$^{-1}$ (\textcolor{red}{CITE!!}). From the work of Hilbert et al. \cite{ray_tracing}, high resolution maps of the ray-traced shear and convergence of galaxy sources at $z=1.3857$ are publicly available. From these maps, an estimate of the `true' lensing of background galaxy ellipticities when traveling through the foreground mass of the Millennium Simulation can be calculated using Equation \eqref{lensing}.

@@

To calculate the convergence and shear contribution of each halo, the physical distance from the halo to the line of sight was needed. To increase the speed of distance calculations, the foreground halo redshifts were first binned to a grid of 100 equally-sized redshift bins $\Delta z\approx0.013857$ from $z=0$ to the sourced redshift of $z=1.3857$, and then converted to physical distances using the cosmology defined by the Millennium Simulation.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Calculating the Smooth-Component Correction} \label{calc_scc}

@@

As discussed in Section \ref{scc}, the halo model does not account for additional mass structures such as filaments. To apply the smooth-component correction to \texttt{Pangloss}, the mean mass density $\rho(z)$ and halo density $\rho_h(z)$ had to be calculated at each of the 100 redshift bins. For $\rho_h(z)$, this was done by summing all halo masses $M_{h,i}$ in a given redshift bin and integrating Equation \eqref{pvolume_element} from ${z-\Delta z/2}$ to ${z+\Delta z/2}$, except for the boundary bins. Once $\rho_h(z)$ had been calculated for each redshift slice, the projected surface mass density was found by integrating over the redshift interval,
\begin{equation}\label{surface_density} 
\sum(z)=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho_s(z)\,dz=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho(z)-\rho_h(z)\,dz,
\end{equation}

\noindent and the convergence contribution of the mass sheet was calculated using Equation \eqref{convergence} with ${z_l=z}$ and ${z_s=1.3857}$. The corrected total convergence for a given background source is simply the sum of the halo contributions and the asfd. There is no additional shear produced by a uniform sheet of mass (citation!), but the corrected reduced shear is now calculated using the corrected total $\kappa$ in Equation \eqref{reduced_shear}. A diagram of this process is shown in Figure \ref{fig:scc}.

The volume and projected surface mass densities of each redshift slice are shown in Figures \ref{fig:volume_densities} and \ref{fig:surface_densities} respectively. (Describe plots)

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Surface and Volume densities for smooth-component correction redshift slices
% \begin{figure*}
%     \centering
%     \begin{subfigure}[H]{0.415\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/scc_vol_den.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{fig:volume_densities}
%     \end{subfigure}
%     ~
%     %% Einstein Ring Figure
%     \begin{subfigure}[H]{0.425\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/scc_surf_deng.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{fig:surface_densities}
%     \end{subfigure}
%     \caption{}
% \end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------
